Leveraging the VTA-TVM Hardware-Software Stack \\for FPGA Acceleration of 8-bit ResNet-18 Inference

### 2018-06-13 dissemination.publication/9375838469ad4029/paper.tex
edits

\paragraph{Runtime and JIT Compilation}
After the TVM optimization and lowering phases, we end up with calls into the accelerator's runtime.
The runtime assembles the accelerator instruction stream and micro-kernels on the fly in a JIT fashion.
This offloads the complexities of generating assembly away from the TVM stack.
The runtime also handles synchronization between the CPU and the FPGA when computation is being offloaded onto the accelerator.

@@

\paragraph{Vanilla Tensor Accelerator (VTA)}
We present VTA, a generic, modular and customizable TPU-like hardware accelerator design.
At its core, the VTA design can perform a single matrix-matrix multiplication per cycle.
These tensorized operations take advantage of the high operational intensity of deep learning workloads.
VTA is programmed with CISC instructions that describe high-level tensor operations, and RISC micro-kernel that dictate low-level data access patterns for CISC compute instructions.
In order to maximize compute resources utilization, VTA relies on a decoupled access-execute pipeline~\cite{Smith:DAE} where memory load, stores and computation can occur concurrently.
We show a high-level schematic of the VTA design in Figure~\ref{fig:vta_overview}: each pipeline stage communicates via dependence queues that guarantee correctness of execution.
Those dependence queues are explicitly programmed via synchronization operations that are added to the instruction stream in the JIT compilation phase.
These synchronization operations are derived from thread context information that are specified by the TVM compiler.
The process is described in more depth in ~\cite{chen:TVM}.

@@

We implement the VTA design on a low-power PYNQ board which incorporates an ARM Cortex A9 dual core CPU clocked at 667MHz and an Artix-7 based FPGA fabric.
On the modest FPGA resources, we implement a $16\times16$ matrix-vector unit clocked at 100MHz that performs products of 8-bit values and accumulates them into a 32-bit register every cycle.
The theoretical peak throughput of this flavor of the VTA design lies around 51GOPS/s.
We allocate 16kB of resources for microkernel cache, 32kB of resources for activation storage, 256kB for parameter storage, and 128kB for the register file (i.e. accumulator storage).
These on-chip buffers are nowhere near large enough to provide enough on-chip storage for a single layer of ResNet, and therefore enable a case study on effective memory reuse and memory access latency hiding.

We offload all convolution layers of the ResNet-18 network onto the FPGA accelerator, except for the first convolution layer C1, which is evaluated on the CPU.
Activations and batch normalization operators are evaluated on the FPGA, while max pooling and fully connected layers are evaluated on the CPU.

@@

\paragraph{Single Kernel Evaluation}
We then evaluate each layer individually to evaluate TVM's ability to efficiently utilize the accelerator's resources.
The results are shown in the Roofline diagram in Figure ~\ref{fig:fpga_latencyhiding}.
Roofline performance diagrams provide insight on how well computation and memory resources are utilized on a given system for different workloads.
Overall, virtual threading in TVM improves performance on all ResNet layers by exploiting latency hiding mechanisms in VTA's decoupled access-execute pipeline.
Peak compute utilization increases from 70\% with no latency hiding to 88\% with latency hiding turned on.

\paragraph{End-to-end ResNet Evaluation}
We leverage TVM to generate ResNet inference kernels on the PYNQ platform and offload as many layers as possible to VTA.
We utilize TVM to generate both schedules for the CPU only and CPU+FPGA implementation.
Due to its shallow convolution depth, the first ResNet convolution layer could not be efficiently offloaded on the FPGA and is instead computed on the CPU.
All other convolution layers in ResNet, however, are amenable to efficient offloading.
Operations including residual layers and max pooling are also performed on the CPU since TVM does not support these operations.

>>>>

\paragraph{Runtime and JIT Compilation}
After the TVM optimization and lowering phases, we obtain a lowered LLVM-compilable program with calls into the accelerator's runtime.
The runtime assembles the accelerator instruction stream and micro-kernels on the fly in JIT fashion.
This offloads the complexities of generating assembly away from the TVM stack.
The runtime also handles synchronization between the CPU and the FPGA when computation is being offloaded onto the accelerator.

@@

\paragraph{Vanilla Tensor Accelerator (VTA)}
We provide an overview of VTA: a generic, modular and customizable TPU-like hardware accelerator design.
At its core, the VTA design can perform a single matrix-matrix multiplication per cycle.
These \emph{tensorized} operations are tailored to take advantage of the high operational intensity of deep learning workloads.
VTA is programmed with CISC instructions that describe high-level, multi-cycle tensor operations, and RISC micro-kernel instructions that dictate low-level data access patterns for CISC compute-instructions.
In order to maximize compute resources utilization, VTA relies on a decoupled access-execute pipeline~\cite{Smith:DAE} where memory load, stores and computation can occur independently, and concurrently.
We show a high-level schematic of the VTA design in Figure~\ref{fig:vta_overview}: each pipeline stage communicates via dependence queues that guarantee correctness of execution.
Those dependence queues are explicitly programmed via synchronization operations that are added to the instruction stream in the JIT compilation phase.
These synchronization operations are derived from thread context information which TVM produces when the user parallelizes a workload across two virtual threads.
This virtual thread lowering process is described in more depth in ~\cite{chen:TVM}.

@@

We implement the VTA design on a low-power PYNQ board which incorporates an ARM Cortex A9 dual core CPU clocked at 667MHz and an Artix-7 based FPGA fabric.
On the modest FPGA resources, we implement a $16\times16$ matrix-vector unit clocked at 100MHz that performs products of 8-bit values and accumulates them into a 32-bit register every cycle.
The theoretical peak throughput of this flavor of the VTA design lies around 51GOPS/s.
We allocate 16kB of resources for microkernel cache, 32kB of resources for activation storage, 256kB for parameter storage, and 128kB for the register file (i.e. accumulator storage).
These on-chip buffers are nowhere near large enough to provide enough on-chip storage for a single layer of ResNet, and therefore provide motivational case-study for effective memory reuse and memory access latency hiding.

We offload all convolution layers of the ResNet-18 network onto the FPGA accelerator, except for the first convolution layer C1, which is evaluated on the CPU due to its low number of input channels.
Activations and batch normalization operators are evaluated on the FPGA, while max pooling and fully connected layers are evaluated on the CPU.

@@

\paragraph{Single Kernel Evaluation}
We then evaluate each layer individually to evaluate TVM's ability to efficiently utilize the accelerator's resources.
The results are shown in the Roofline diagram in Figure ~\ref{fig:fpga_latencyhiding}.
Roofline performance diagrams provide insight on how well computation and memory resources are utilized on a given system for different workloads.
Overall, virtual threading in TVM improves performance on all ResNet layers by exploiting latency hiding mechanisms in VTA's decoupled access-execute pipeline.
Peak compute utilization increases from 70\% with no virtual threading to 88\% with virtual threading turned on (to hide memory access latency).

\paragraph{End-to-end ResNet Evaluation}
We leverage TVM to generate ResNet inference kernels on the PYNQ platform and offload as many layers as possible to VTA.
We utilize TVM to generate both schedules for the CPU only and CPU+FPGA implementation.
Due to its shallow convolution depth, the first ResNet convolution layer could not be efficiently offloaded on the FPGA and is instead computed on the CPU.
All other convolution layers in ResNet, however, are amenable to efficient offloading.
Operations including residual layers and max pooling are also performed on the CPU since our TVM extensions do not support these operations yet.
